---
title: "Initial Korean covid 19"
author: "Jovygomez"
date: "October 12, 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was un coded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

Question- Is dying independent of sex?

$H_0$ There is no diffrence in dying depending on sex.  
$A_0$ There is a diffrence in dying depending on sex. 


result~sex
categorical~categorical. 
We will use the helper file CAT~CAT


## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")
```

The bar chart shows that more males died than female. This shows that dying dependent on sex .  


## Numerical results

```{r}
table1 <- xtabs(~result+ Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results shows that amoung the people who dies was 64% of males who died and 36% of females who died. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The infrential results shows that the chi-squared test has a p-value of 0.001, and the the Fishers exact test has a p-value of 0.00009 because both of the p-values are under the acceptable rate of 0.05 we will reject the null hypothesis. The Fishers Exact test also states that men are 2.3 more times to die of Covid19 in Korea than women are

## Conclusions

